---
title: "Coursera Capstone Milestone Report"
author: "Glenn Kerbein"
date: "July 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Report Details

The goal of this project is just to display that you’ve gotten used to working with the data and that you are on track to create your prediction algorithm. Please submit a report on [R Pubs](http://rpubs.com/) that explains your exploratory analysis and your goals for the eventual app and algorithm. This document should be concise and explain only the major features of the data you have identified and briefly summarize your plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager. You should make use of tables and plots to illustrate important summaries of the data set.

The motivation for this project is to:
1. Demonstrate that you've downloaded the data and have successfully loaded it in.
2. Create a basic report of summary statistics about the data sets.
3. Report any interesting findings that you amassed so far.
4. Get feedback on your plans for creating a prediction algorithm and Shiny app.

## Demonstrate successful data fetch and load

Download the SwiftKey dataset. If it does not exist, then download and unzip it.
```{r echo=TRUE}
SwiftKey_data_url = "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
SwiftKey_destination_zip = "Coursera-SwiftKey.zip"

if (!file.exists("./data/final")) {
  if (!file.exists(paste0("./data/", SwiftKey_destination_zip))) {
    download.file(
      SwiftKey_data_url,
      destfile = SwiftKey_destination_zip,
      quiet=TRUE
    )
  }
  
  unzip("Coursera-SwiftKey.zip", exdir = "./data");
}
```

According to the specifications, load the English data set only.
```{r echo=TRUE}
blogs <- readLines("./data/final/en_US/en_US.blogs.txt", encoding = "UTF-8", skipNul=TRUE)
twitter <-  readLines("./data/final/en_US/en_US.twitter.txt", encoding = "UTF-8", skipNul=TRUE)

# News dataset looks like a raw binary file to R. To mitigate this, open the file as a stream,
# then read the data, line by line, into a new variable.
con <- file("./data/final/en_US/en_US.news.txt", open="rb")
news <- readLines(con, encoding="UTF-8", skipNul=TRUE)
close(con)
rm(con)
```

Grab sample data.

```{r echo=TRUE}
head(blogs, 10)
```
```{r echo=TRUE}
head(twitter, 5)
```
```{r echo=TRUE}
head(news, 5)
```

## Basic report of summary statistics about the data sets


## Report any interesting findings that you amassed so far


## Get feedback on your plans for creating a prediction algorithm and Shiny app

